Donor Responses to Civil Society Repression

### 2024-11-01 manuscript.qmd
Super minor edits

Though the majority of anti-NGO laws have targeted foreign funding, other laws seek to impose barriers to entry, such as strict and burdensome registration requirements, and barriers to advocacy that explicitly limit NGO political participation [@Chaudhry:2022]. Donor countries respond differently to these types of oppression. @ChaudhryHeiss:2024 (see also @ChaudhryHeiss:2022a) find that bilateral donors are quite responsive to barriers to advocacy—on average, total foreign aid decreases by 30–40% (or \$100 in a typical country) following the passage of an advocacy-focused NGO law. @RightSpringmanWibbels:2024 find similar and more nuanced trends in donor responses: donors committed to political advocacy   (i.e. those that fund democracy and civil society promotion activities) reduce funding for advocacy programs by more than 70% in response to new restrictions, and the reduction in aid persists for several years.

Overall bilateral aid does not decrease significantly in response to new barriers to entry, but the distribution channels of that aid shift as donor countries channel more aid through domestic rather than foreign NGOs [@ChaudhryHeiss:2024]. USAID even explicitly encourages its program managers to rely more on domestic, in-country NGOs when facing a restrictive legal environment [@USAID:2014, 6–7]. In addition to changing the amounts and recipients of foreign aid, donor agencies often adjust the focus of the programs they fund, redirecting funding away from programs related to politically sensitive issues like human rights, anti-corruption, and democratic development and towards tamer topics that are more compatible with recipient-country preferences, such as education, health, and humanitarian services [@Bush:2015; @CheesemanDesrosiers:2023].

>>>>

Though the majority of anti-NGO laws have targeted foreign funding, other laws seek to impose barriers to entry, such as strict and burdensome registration requirements, and barriers to advocacy that explicitly limit NGO political participation [@Chaudhry:2022]. Donor countries respond differently to these types of oppression. @ChaudhryHeiss:2024 find that bilateral donors are quite responsive to barriers to advocacy—on average, total foreign aid decreases by 30–40% (or \$100 in a typical country) following the passage of an advocacy-focused NGO law [see also @ChaudhryHeiss:2022a]. @RightSpringmanWibbels:2024 find similar and more nuanced trends in donor responses: donors committed to political advocacy   (i.e. those that fund democracy and civil society promotion activities) reduce funding for advocacy programs by more than 70% in response to new restrictions, and the reduction in aid persists for several years.

Overall bilateral aid does not decrease significantly in response to new barriers to entry, but the distribution channels of that aid shift as donor countries channel more aid through domestic rather than foreign NGOs [@ChaudhryHeiss:2024]. USAID even explicitly encourages its program managers to rely more on domestic, in-country NGOs when facing a restrictive legal environment [@USAID:2014]. In addition to changing the amounts and recipients of foreign aid, donor agencies often adjust the focus of the programs they fund, redirecting funding away from programs related to politically sensitive issues like human rights, anti-corruption, and democratic development and towards tamer topics that are more compatible with recipient-country preferences, such as education, health, and humanitarian services [@Bush:2015; @CheesemanDesrosiers:2023].
